\chapter{Acknowledgments}


Embarking on this thesis journey would have been a considerably more challenging task without the support and guidance of numerous individuals to whom I owe my deepest gratitude.

First, I express my heartfelt appreciation to Philipp Schröeppel, my advisor. His constant encouragement, insightful feedback, and unwavering faith in my abilities were pivotal in shaping this work. Your mentorship, Philipp, has not only benefitted this thesis but also played a profound role in my personal and academic growth.

I am equally grateful to Prof. Thomas Brox for introducing me to this topic and entrusting me with its exploration. His flexibility and consistent availability, even during the busiest times, provided a reassuring anchor throughout the course of this project. Your dedication to students and passion for the subject have been truly inspiring.

To my sister Shraddha Barke and my mother Alka Barke, your enduring love, patience, and belief in me have been my strength during the highs and lows of this journey. The countless sacrifices you have made and your unwavering support have been the wind under my wings. You reminded me of the larger picture when I lost sight of it and celebrated with me during the milestones. This achievement is as much yours as mine. I would also like to thank my late father Govind Barke, who stood by my decision to quit my old job and pursue higher education in Germany. You are the rock that I stand on. 

Lastly, to all those who have supported me directly or indirectly, shared their invaluable insights, or believed in me, I extend my sincere thanks. The culmination of this thesis is a testament to the collective effort of all those mentioned and many others who stood by me during this phase of my life.